Quantile Trend Filtering via Spingarn's Method

### 2017-05-15 Notes/detrendify_spingarn.tex
Small edits to clean up.
typos
Added some references; fixed some typos.
Added some notes on detrending versus denoising and detrending.

Examples:
\begin{itemize}
\item Air quality 
\item ECG \cite{Sanyal2012}
\item Chromatogram baseline estimation  \cite{Ning2014}
\end{itemize}

Things to do:
\begin{itemize}
	\item Convergence of the algorithm
	\item Convergence rate?
	\item Timing experiments of LP versus Spingarn
	\item Make an R package - detrendr
	\item Compare on synthetic data quality of solution with existing methods
	\item Do comparisons on real data examples
\end{itemize}

@@

\begin{eqnarray*}
\text{minimize}\; && f_1(\V{\theta}) + f_2(\V{\eta}) \\
\text{subject to}\; && \V{\eta} = \M{D}\V{\theta}
\end{eqnarray*}
where
\begin{eqnarray*}
f_1(\V{\theta}) = \sum_{i=1}^n \rho_\tau(\VE{y}{i} - \VE{\theta}{i}) \quad\quad \text{and} \quad\quad
f_2(\V{\eta}) = \lVert \V{\eta} \rVert_1.
\end{eqnarray*}

@@

\begin{eqnarray*}
\left [\prox{tf_1}{\V{\theta}} \right ]_i & = & \prox{(t/n)\rho_\tau}{\VE{y}{i} - \VE{\theta}{i}}, \\
\left [\prox{tf_2}{\V{\eta}} \right ]_j & = & S(\VE{\eta}{j}, t\lambda).
\end{eqnarray*}

@@

banded linear system solve, with bandwidth $k + 1$.
This linear solve can be accomplished in $\mathcal{O}(n(k + 1)^2)$. The first solve using a banded Cholesky decomposition requires $\mathcal{O}(n(k + 1)^2)$. Subsequent solves require
$\mathcal{O}(n(k+1))$.

@@

The total cost when $n \gg p$ is $\mathcal{O}(nk)$.

@@

%\section{Problem}
%
%\begin{eqnarray*}
%\text{minimize}\; \left [\sum_{i=1}^n w_i h\left (\langle \V{x}_i, \V{\beta} \rangle \right) \right ] + \lambda \lVert \V{\beta} \rVert_1,
%\end{eqnarray*}
%where
%\begin{eqnarray*}
%h(r) & = & \left (1 - r \right )_+ \\
%\partial h(r) & = & \begin{cases}
%-1 & r < 1 \\
%[-1,0] & r = 1 \\
%0 & r > 1.
%\end{cases}
%\end{eqnarray*}

%\section*{ADMM formulation}
%
%\begin{eqnarray*}
%\underset{\V{\beta},\V{z}}{\text{minimize}}\;  & & \left [\sum_{i=1}^n w_i h\left (\VE{z}{i} \right) \right ] + \lambda \lVert \V{\beta} \rVert_1 \\
%\text{subject to}\; & & \V{z} - \M{X}\V{\beta} = \V{0}.
%\end{eqnarray*}
%
%ADMM Updates
%
%\begin{eqnarray*}
%\Vn{z}{m+1} & = & \underset{\V{z}}{\arg\min}\;  \left [\sum_{i=1}^n w_i h\left (\VE{z}{i} \right) \right ] + \frac{\rho}{2} \lVert \V{z} - \M{X}\Vn{\beta}{m} + \Vn{u}{m} \rVert_2^2 \\
%\VnE{z}{m+1}{i} & = & \prox{(w_i/\rho) h}{\langle \V{x}_i, \Vn{\beta}{m} \rangle - \VnE{u}{m}{i}} \\
%\Vn{\beta}{m+1} & = & \underset{\V{\beta}}{\arg\min}\;  \frac{\lambda}{\rho} \lVert \V{\beta} \rVert_1 + \frac{1}{2} \lVert \Vn{z}{m+1} - \M{X}\V{\beta} + \Vn{u}{m} \rVert_2^2 \\
%\Vn{u}{m+1} & = & \Vn{u}{m} + \Vn{z}{m+1} - \M{X}\Vn{\beta}{m+1}.
%\end{eqnarray*}
%
%
%\begin{eqnarray*}
%\text{minimize}\; && f_1(\V{\beta}) + f_2(\M{X}\V{\beta}) \\
%\end{eqnarray*}

>>>>

Examples:
\begin{itemize}
\item Air quality 
\item ECG \cite{Sanyal2012}
\item Chromatogram baseline estimation  \cite{Ning2014, Ilewicz2015}
\item Galaxy spectrum baseline estimation \cite{Ilewicz2015,Bacher2016}
\end{itemize}

Need to decide between detrending versus detrending + denoising. The BEADS \cite{Ning2014} method does both. We may wish to focus on detrending and then use wavelet SURE denoising as a postprocessing step, i.e. do a two-stage procedure, both of which can be done in linear time. On paper this should be faster than the BEADS procedure. Or we may just want to stick with detrending.  There's Matlab code for BEADS, and the BEADS paper also points to two other popular methods in chromotography.

Things to do:
\begin{itemize}
	\item Convergence of the algorithm
	\item Convergence rate \cite{He2012, He2015, Davis2017}
	\item Timing experiments of LP versus Spingarn
	\item Make an R package - detrendr
	\item Compare on synthetic data quality of solution with existing methods
	\item Do comparisons on real data examples
\end{itemize}

@@

\begin{eqnarray*}
\text{minimize}\; && f_1(\V{\theta}) + f_2(\V{\eta}) \\
\text{subject to}\; && \V{\eta} = \M{D}\V{\theta}
\end{eqnarray*}
where
\begin{eqnarray*}
f_1(\V{\theta}) = \frac{1}{n}\sum_{i=1}^n \rho_\tau(\VE{y}{i} - \VE{\theta}{i}) \quad\quad \text{and} \quad\quad
f_2(\V{\eta}) = \lambda\lVert \V{\eta} \rVert_1.
\end{eqnarray*}

@@

\begin{eqnarray*}
\left [\prox{tf_1}{\V{\theta}} \right ]_i & = & \VE{y}{i} - \prox{(t/n)\rho_\tau}{\VE{y}{i} - \VE{\theta}{i}}, \\
\left [\prox{tf_2}{\V{\eta}} \right ]_j & = & S(\VE{\eta}{j}, t\lambda).
\end{eqnarray*}

@@

@@

The total cost is $\mathcal{O}(nk)$.

@@
